### 2024-12-29 skripsi/chapter4.tex
ch4 add scenario 2, todo: predictions for exact burgers

After this, in the third step, the solution functions \(u \) are generated with 20 modes in time and 20 in space and 500 unique function samples for each viscosity value. In total 1500 unique solution functions are generated.

As for the forcing term, it is computed using \lccref{eq:forced_viscous_burgers}. The solution field \(u\) and forcing term \(f\) are modeled with Fourier series in \lccrefs{eq:fourier_field,eq:fourier_force} where \(k\) is a vector of spatial and temporal wave number such that \(k_t\) is the temporal wave number and \(k_x\) is the spatial wave number.
\begin{align}
  u\left(x, t \right)           & = \sum_{k_{x}} \sum_{k_{t}} \hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)} \label{eq:fourier_field}        \\
  f\left(x, t \right)           & = \sum_{k_{x}}\sum_{k_{t}} \hat{f}_k e^{2\pi i(k_{x}x+k_{t}t)} \label{eq:fourier_force}
\end{align}
Substituting the terms in \lccref{eq:forced_viscous_burgers} with the respective Fourier series, we get the equation \lccref{eq:forced_viscous_burgers_fourier_substitution}. A problem one notices is the nonlinear term \(u\pdv{u}{x}\). A naive approach would multiply all wave numbers terms with each other.
\begin{align}
  u\pdv{u}{x} = \left(\sum_{k_{x}} \sum_{k_{t}} \hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)}\right) \times \left(\sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{x})\hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)}\right)
\end{align}
This is computationally expensive operation with \(N^2\) multiplications \autocite{lariosMATH934BURGERS2021,robertsDealiasedConvolutionsPseudospectral2011,shenSpectralMethodsAlgorithms2011,orszagComparisonPseudospectralSpectral1972}. To avoid this complexity, the term is first reformulated into \(\pdv{x}(u^2/2)\). Then to avoid aliasing, the coefficients are padded with 50\% zeros such that the padded coefficients are 3/2 times the size of the original \autocite{orszagEliminationAliasingFiniteDifference1971, lariosMATH934BURGERS2021}. Then we use the padded coefficients to compute the function values in the physical domain using the inverse transform. And then, the point wise squaring operation is performed and transform the results back to spectral domain. The resulting coefficients are then trimmed back to their original size before padding. Finally, we multiply the resulting coefficients \(\hat{uu}_k\) by the derivative constants. This operation is much more efficient with only \(1.5 N\) multiplications. This is still more efficient than the naive approach even after taking into account the transforms involved adds \(O(2\times 1.5Nln(1.5N))\) operations if using a Fast Fourier Transform algorithm. Putting all the above together, results in \lccref{eq:forced_viscous_burgers_fourier_substitution}.
\begin{equation}
  \begin{split}
    \sum_{k_{x}} \sum_{k_{t}} \hat{f}_k e^{2\pi i(k_{x}x+k_{t}t)} =
    & \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{t}) \hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)} \\
    & + \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{x})\hat{uu}_k e^{2\pi i(k_{x}x+k_{t}t)} \\
    & - \nu\sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{x})^2\hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)} \label{eq:forced_viscous_burgers_fourier_substitution}
  \end{split}
\end{equation}
\begin{align}
  \sum_{k_{x}} \sum_{k_{t}} \hat{f}_k & = \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{t}) \hat{u}_k + (2\pi i k_{x})\hat{uu}_k - \nu(2\pi i k_{x})^2\hat{u}_k \label{eq:forced_viscous_burgers_fourier} \\
  \hat{f}_k & = (2\pi i k_{t}) \hat{u}_k + (2\pi i k_{x})\hat{uu}_k - \nu(2\pi i k_{x})^2\hat{u}_k \label{eq:forced_viscous_burgers_coeff}
\end{align}
After simplifying the equation, we get the coefficients as in \lccref{eq:forced_viscous_burgers_coeff}. Using this equation, the exact forcing term corresponding to the randomly generated solution can be computed with relatively low cost. Finally, the function values of both the solutions and forcing terms are perturbed by adding Gaussian noise with a mean of zero and standard deviation of 10\% of the function value standard deviation to the values of the inverse transform. The perturbed coefficients are then recomputed from the sum and the final perturbed functions are obtained.

@@

\subsection{Data Transformation}
\noindent The second phase of the computational model transforms the data that has been prepared into a form that can be used with the proposed model. In this study, there are two ways a dataset can be used. The first and simplest way is to predict the solution function from the forcing term for a given set of PDE parameters. In this case, the input function is the forcing term. In other words, the input function is the predictor or independent variable that will be used to predict the solution function. As such, the output function is the solution function. The antiderivative dataset is used in this way. The derivative, which represents the acceleration function, is used to predict the velocity function. Which means that the input function is the acceleration function and the output function is the velocity function. The second use of a dataset depends on whether it can be interpreted as a mapping of a previous function to the next function. In other words, a series of functions related to the previous function as a sum of the previous function and a perturbation. A commonly used form of this are time series. Generally, a time series can be represented as \lccref{eq:time_step_function}. This function describes the evolution of function \(f\) as a sum of some initial state \(f^0\) and perturbation which is its rate of change \(\dv{f}{t}\) at a point in time close to the current time multiplied by the time step \(\delta t\).
\begin{equation}
  f^{t+1} = f^{t} + \delta t \times \dv{f}{t} \label{eq:time_step_function}
\end{equation}
In this case the previous function and any other independent variables like the forcing term involved in the time derivative \(\dv{f}{t}\) is used to predict the perturbation which is the time derivative itself. It would mean that each sample function in the data would need to at least have two dimensions. This is because one dimension represents the evolution of the function over time, while the other dimensions define the spatial domain where the function's values are computed at each time step. This condition is true for both the Burgers' equation dataset and the WeatherBench 2 collection of ERA5 datasets. First, the Burgers' equation dataset, while is also able to be used in the first way as a mapping between the forcing term and the solution, one of the traditional ways of solving it is using the method of lines in conjunction with another method such as the finite differences method or the spectral method. In both of these, the Burgers' equation is formulated as an initial value problem, using the finite differences to computing the time derivative. % TODO: add figures and citation for this, or if this is avaliable in a different place in the thesis a refference to that.
The method of lines approach in one of its forms is equal to \lccref{eq:time_step_function} \autocite{schiesserNumericalMethodLines2012,sadikuSimpleIntroductionMethod2000}. As such the Burgers' equation dataset can also be used in this way. In this case, algebraic manipulation of \lccref{eq:forced_viscous_burgers} results in the time derivative of the solution as in \lccref{eq:forced_viscous_burgers_time_derivative}.
\begin{align}
  \pdv{u}{t} = f - u\pdv{u}{x} + \nu\pdv[2]{u}{x} \label{eq:forced_viscous_burgers_time_derivative}
\end{align}
As can be seen, the time derivative is expressed in terms of the solution function, the forcing term and any other partial derivatives. Because of this, it means we can train the model to learn the relationship between the solution function combined with the forcing term and the time derivative. Putting this together with \lccref{eq:time_step_function}, it is possible to predict the solution function in a future time based on the current solution function and forcing term. The input functions in this case are both the value of the solution function and the forcing term at the current time step. The output function is the time derivative which is computed by rearranging \lccref{eq:time_step_function} resulting in \lccref{eq:time_derivative_time_step_function}.
\begin{equation}
  \dv{f}{t} = \frac{f^{t+1} - f^{t}}{\delta t} \label{eq:time_derivative_time_step_function}
\end{equation}
Perhaps in a more obvious manner, the same is also applied to the WeatherBench 2 collection of ERA5 datasets. The 2-meter temperature is a series of snapshots in time of the distribution of temperature all over the world at 2 meters above the surface. The input function in this case is the temperature distribution at the current time step. The output function is the time derivative of the temperature distribution computed using \lccref{eq:time_derivative_time_step_function}.

@@

The training can proceed to solve \lccref{eq:lssvr_solution}. As the original LSSVR equation was made for regression of a single value of \(\vb{y} \), we constructed an extended the original equation in order to train multiple regression models to predict each coefficient in an output function sample. The assumption of this extension is that the inputs and model hyperparameters such as kernels and the regularization constant \(C \) are the same across individual regressors. Therefore, it follows that the solution vector and target vector can both be transformed into their matrix form with each column representing the individual regressor which results in \lccref{eq:lssvr_solution_vectorized}. This vectorized version is equal to \(m \) separate instances of \lccref{eq:lssvr_solution}. In essence, instead of a vector of scalar sample output values and a vector of model parameters, the extended equation consists of a matrix of rows of vector output values and a matrix of model parameters with each column representing the separate model for each output and row representing the Lagrange multipliers for each sample. The procedure to solve \lccref{eq:lssvr_solution_vectorized} adjusts \lccref{alg:lssvr_training} for the extra dimension spanning the multiple outputs of size \(m\).
\begin{equation} \label{eq:lssvr_solution_vectorized}
  \begin{bmatrix}
    0          & \vb{1}_{n}^{\intercal} \\
    \vb{1}_{n} & \Omega + \frac{I}{C}
  \end{bmatrix}
  \begin{bmatrix}
    b_0           & \cdots & b_m \\
    \vb{\alpha}_0 & \cdots & \vb{\alpha}_m\\
  \end{bmatrix}
  =
  \begin{bmatrix}
    0        & \cdots & 0 \\
    \vb{y}_0 & \cdots & \vb{y}_m \\
  \end{bmatrix}
\end{equation}

@@

\noindent The utilities module provide common tools that are used when working with the developed software. First, because the basis function used in this study is the Fourier basis, a large proportion of concern is the storage of Fourier basis coefficients. The Fourier series representation that is used in this study is the complex representation as show in \lccref{eq:fourier_series}. However, as previously mentioned, least squares support vector machines is constructed to work with real numbers. In order for LSSVR to learn the relationship between complex numbers, a representation of complex numbers in terms of real numbers is necessary. Since sample features and labels for LSSVR need to be flattened into one dimension which result in a matrix with rows of samples, the representation of complex numbers using real numbers can be done after features and labels have been flattened. An example of such a matrix of two samples with four features each can be seen in \lccref{eq:complex_matrix}. To represent the complex number features as real numbers, once can simply split a single complex number into two real numbers. When this is done to the entire matrix of complex numbers, the result is \lccref{eq:complex_matrix_real_rep}.
\begin{align}
  \begin{bmatrix}
    -1.03+0i & -0.52+0.39i & -0.54+0i & -0.52-0.39i\\
    -0.85+0i & -0.39+0.19i &  0.87+0i & -0.39-0.19i
  \end{bmatrix}\label{eq:complex_matrix}\\
  \begin{bmatrix}
    -1.03 & 0 & -0.52 & 0.39 & -0.54 & 0 & -0.52 & -0.39\\
    -0.85 & 0 & -0.39 & 0.19 &  0.87 & 0 & -0.39 & -0.19
  \end{bmatrix}\label{eq:complex_matrix_real_rep}
\end{align}
There are several ways to do this in Pytorch, such as extracting the real and imaginary element components into separate matrices and then interleaving the columns to create the real representation. However, a more performant version can be made by using built-in functions of the library that directly allows views of complex valued tensors as real tensors and vice versa. The \verb|view_as_real| and \verb|view_as_complex| functions allow the conversion between real and complex representations by representing complex valued tensors by adding a dimension of size two for each component of complex numbers. The resulting implementation can be seen in \lccref{fig:complex_matrix_conversion_implementation}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
def to_real_coeff(x: torch.Tensor) -> torch.Tensor:
    return torch.view_as_real(x).flatten(-2)

@@

\begin{table}[H]
  \caption{The configuration of experimental scenarios in this study.} \label{table:experimental_scenarios}
  \centering
  \begin{tabularx}{\textwidth}{ L{0.75} *{3}{L{1.09}} } % compute these using approx barycentric coords (x + 3y = 4, x = 0.8y)
    \toprule
    Configuration & Scenario 1 & Scenario 2 & Scenario 3       \\
    \midrule
    Data Source & Synthetic (computed from equations) & Synthetic (computed from equations) with three viscosity values & ERA5 dataset \\\addlinespace[0.5em]
    Data Augmentations & Perturbations of three different noise levels & Perturbation of 10\% the mean standard deviation of function values &  \\\addlinespace[0.5em]
    Data Sample Size & 5000 unique functions for each noise level & 500 unique functions for each viscosity value & 7000 time steps \\\addlinespace[0.5em]
    Number of Modes &  &  &  \\\addlinespace[0.5em]
    Scaling & Standard scaling & Standard scaling & Standard scaling\\\addlinespace[0.5em]
    Dataset Split & 80\% Training and 20\% Testing for each noise level & 80\% Training and 20\% Testing for each viscosity value & 80\% Training and 20\% Testing \\\addlinespace[0.5em]
    Regularization Parameter (\(C\)) & 1.0 & 1.0 & 1.0 \\\addlinespace[0.5em]
    Kernel & RBF (\(\sigma = features \times X.var()\)) & RBF (\(\sigma = features \times X.var()\)) & RBF (\(\sigma = features \times X.var()\)) \\\addlinespace[0.5em]
    Additional Tests & Exact solutions & Exact solutions &  \\\addlinespace[0.5em]
    Interpretation & p-matrix and correlation image & p-matrix and correlation image & p-matrix and correlation image \\
    \bottomrule
  \end{tabularx}
\end{table}

Each scenario serves a different purpose. The first scenario is constructed to establish whether the model can learn any relationships at all in data that implicitly defines differential equations. This scenario is a proof of concept of whether the model is capable of accurately learning the relationship between derivatives and their anti-derivatives. As the simplest case, it provides a foundation for understanding the model and its behavior. The 5\% noise model is then used to predict the antiderivative of a cosine function with a period of one as shown in \lccref{eq:sine_derivative}. The exact solution is known as \lccref{eq:sine_function} \autocite{abramowitzHandbookMathematicalFunctions1972}. The second goal of this scenario is to determine how noise affects the model's predictions. This is especially important for the inverse problem because of how sensitive the derivative is to changes in the antiderivative. To further solidify how the model fares for inverse problems, this scenario will also assess the performance of a trained model for a very simple inverse problem setup to determine the derivatives of the testing subset from the antiderivatives. This setup will utilize the model as a surrogate forward solver and optimize for a loss function using the inverse function of the SpectralSVR class as described in \lccref{sec:impl_model}.
\begin{align}
  f(x) &= 2\pi cos\left(2\pi x\right)\label{eq:sine_derivative}\\
  f(x) &= sin\left(2\pi x\right)\label{eq:sine_function}
\end{align}

@@

\subsection{Scenario 2}
\noindent The second scenario results are presented in two parts which are the coefficient predictions and function value predictions.

>>>>

After this, in the third step, the solution functions \(u \) are generated with 25 modes in time and 25 in space and 40 unique function samples for each viscosity value. In total 120 unique solution functions are generated. To ensure the same solutions are generated every time, the random number generator uses a new instance of the generator with the same seed number. This results in the solutions being the same for all viscosity and other parameters of the equation are adapted to this.

As for the forcing term, it is computed using \lccref{eq:forced_viscous_burgers}. The solution field \(u\) and forcing term \(f\) are modeled with Fourier series in \lccrefs{eq:fourier_field,eq:fourier_force} where \(k\) is a vector of spatial and temporal wave number such that \(k_t\) is the temporal wave number and \(k_x\) is the spatial wave number.
\begin{align}
  u\left(x, t \right) & = \sum_{k_{x}} \sum_{k_{t}} \hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)} \label{eq:fourier_field} \\
  f\left(x, t \right) & = \sum_{k_{x}}\sum_{k_{t}} \hat{f}_k e^{2\pi i(k_{x}x+k_{t}t)} \label{eq:fourier_force}
\end{align}
Substituting the terms in \lccref{eq:forced_viscous_burgers} with the respective Fourier series, we get the equation \lccref{eq:forced_viscous_burgers_fourier_substitution}. A problem one notices is the nonlinear term \(u\pdv{u}{x}\). A naive approach would multiply all wave numbers terms with each other.
\begin{align}
  u\pdv{u}{x} = \left(\sum_{k_{x}} \sum_{k_{t}} \hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)}\right) \times \left(\sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{x})\hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)}\right)
\end{align}
This is computationally expensive operation with \(N^2\) multiplications \autocite{lariosMATH934BURGERS2021,robertsDealiasedConvolutionsPseudospectral2011,shenSpectralMethodsAlgorithms2011,orszagComparisonPseudospectralSpectral1972}. To avoid this complexity, the term is first reformulated into \(\pdv{x}(u^2/2)\). Then to avoid aliasing, the coefficients are padded with 50\% zeros such that the padded coefficients are 3/2 times the size of the original \autocite{orszagEliminationAliasingFiniteDifference1971, lariosMATH934BURGERS2021}. Then we use the padded coefficients to compute the function values in the physical domain using the inverse transform. And then, the point wise squaring operation is performed and transform the results back to spectral domain. The resulting coefficients are then trimmed back to their original size before padding. Finally, we multiply the resulting coefficients \(\hat{uu}_k\) by the derivative constants. This operation is much more efficient with only \(1.5 N\) multiplications. This is still more efficient than the naive approach even after taking into account the transforms involved adds \(O(2\times 1.5Nln(1.5N))\) operations if using a Fast Fourier Transform algorithm. Putting all the above together, results in \lccref{eq:forced_viscous_burgers_fourier_substitution}.
\begin{equation}
  \begin{split}
    \sum_{k_{x}} \sum_{k_{t}} \hat{f}_k e^{2\pi i(k_{x}x+k_{t}t)} =
     & \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{t}) \hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)}                                                              \\
     & + \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{x})\hat{uu}_k e^{2\pi i(k_{x}x+k_{t}t)}                                                            \\
     & - \nu\sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{x})^2\hat{u}_k e^{2\pi i(k_{x}x+k_{t}t)} \label{eq:forced_viscous_burgers_fourier_substitution}
  \end{split}
\end{equation}
\begin{align}
  \sum_{k_{x}} \sum_{k_{t}} \hat{f}_k & = \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{t}) \hat{u}_k + (2\pi i k_{x})\hat{uu}_k - \nu(2\pi i k_{x})^2\hat{u}_k \label{eq:forced_viscous_burgers_fourier} \\
  \hat{f}_k                           & = (2\pi i k_{t}) \hat{u}_k + (2\pi i k_{x})\hat{uu}_k - \nu(2\pi i k_{x})^2\hat{u}_k \label{eq:forced_viscous_burgers_coeff}
\end{align}
After simplifying the equation, we get the coefficients as in \lccref{eq:forced_viscous_burgers_coeff}. Using this equation, the exact forcing term corresponding to the randomly generated solution can be computed with relatively low cost. Finally, the function values of both the solutions and forcing terms are perturbed by adding Gaussian noise with a mean of zero and standard deviation of 5\% of the function value standard deviation to the values of the inverse transform. The perturbed coefficients are then recomputed from the sum and the final perturbed functions are obtained.

An example of the solution and forcing term pairs that has been generated is shown in \lccref{fig:burgers_data}. Using the same seed we can see that the solution functions of the same sample are the same for the most part. However, with the forcing terms, we see much larger amplitudes for higher frequencies for larger viscosity values. This is inline with \lccref{eq:forced_viscous_burgers_coeff}. The squared multiplier gets large much more quickly for higher frequencies due to the squaring operation. The since this effect is controlled by the viscosity via multiplication, the higher the viscosity, the larger the high frequency amplitudes of the forcing term. This in other words mean that more viscous fluids require quadratically more force to affect the solution in the same way with lower viscosities.

\begin{figure}[H]
  \centering
  \begin{subfigure}{0.49\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/burgers_solution_0.0.pgf}
    \end{adjustbox}
    \caption{Solution function for \(\nu=0\).}\label{fig:burgers_solution_0.0}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/burgers_forcing_0.0.pgf}
    \end{adjustbox}
    \caption{Forcing function for \(\nu=0\).}\label{fig:burgers_forcing_0.0}
  \end{subfigure}
  % \\[-0.7\baselineskip]
  \begin{subfigure}{0.49\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/burgers_solution_0.01.pgf}
    \end{adjustbox}
    \caption{Solution function for \(\nu=0.01\).}\label{fig:burgers_solution_0.01}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/burgers_forcing_0.01.pgf}
    \end{adjustbox}
    \caption{Forcing function for \(\nu=0.01\).}\label{fig:burgers_forcing_0.01}
  \end{subfigure}
  % \\[-0.7\baselineskip]
  \begin{subfigure}{0.49\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/burgers_solution_0.1.pgf}
    \end{adjustbox}
    \caption{Solution function for \(\nu=0.1\).}\label{fig:burgers_solution_0.1}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/burgers_forcing_0.1.pgf}
    \end{adjustbox}
    \caption{Forcing function for \(\nu=0.1\).}\label{fig:burgers_forcing_0.1}
  \end{subfigure}
  \caption{Example sample pairs of solution and forcing terms for the Burgers' equation dataset.}\label{fig:burgers_data}
\end{figure}

@@

\subsection{Data Transformation}\label{sec:data_transformation}
\noindent The second phase of the computational model transforms the data that has been prepared into a form that can be used with the proposed model. In this study, there are two ways a dataset can be used. The first and simplest way is to predict the solution function from the forcing term for a given set of PDE parameters. In this case, the input function is the forcing term. In other words, the input function is the predictor or independent variable that will be used to predict the solution function. As such, the output function is the solution function. The antiderivative dataset is used in this way. The derivative, which represents the acceleration function, is used to predict the velocity function. Which means that the input function is the acceleration function and the output function is the velocity function.

The second use of a dataset depends on whether it can be interpreted as a mapping of a previous function to the next function. In other words, a series of functions related to the previous function as a sum of the previous function and a perturbation. A commonly used form of this are time series. Generally, a time series can be represented as \lccref{eq:time_step_function}. This function describes the evolution of function \(f\) as a sum of some initial state \(f^0\) and perturbation which is its rate of change \(\dv{f}{t}\) at a point in time close to the current time multiplied by the time step \(\delta t\). This is the first order Taylor polynomial. Adding more terms with higher order derivatives can be done to increase the accuracy. This describes the same process that is used in the finite differences method.
\begin{equation}
  f^{t+1} = f^{t} + \delta t \times \dv{f}{t} + \dots \label{eq:time_step_function}
\end{equation}
In this case the previous function and any other independent variables like the forcing term involved in the right-hand side of the equation is used to predict the difference between the future state and the current state. It would mean that each sample function in the data would need to at least have two dimensions. This is because one dimension represents the evolution of the function over time, while the other dimensions define the spatial domain where the function's values are computed at each time step. This condition is true for both the Burgers' equation dataset and the WeatherBench 2 collection of ERA5 datasets. First, the Burgers' equation dataset, while is also able to be used in the first way as a mapping between the forcing term and the solution, one of the traditional ways of solving it is using the method of lines. This entails solving the dimensions other that time with methods such as the finite differences method or the spectral method. The time dimension is then separately solved using the finite differences method such as the form shown in \lccref{eq:time_step_function} \autocite{schiesserNumericalMethodLines2012,sadikuSimpleIntroductionMethod2000}. This is the form that the Burgers' equation dataset will be used in. In this case, algebraic manipulation of \lccref{eq:forced_viscous_burgers} results in the time derivative of the solution as in \lccref{eq:forced_viscous_burgers_time_derivative}.
\begin{align}
  \pdv{u}{t} = f - u\pdv{u}{x} + \nu\pdv[2]{u}{x} \label{eq:forced_viscous_burgers_time_derivative}
\end{align}
The most important component of the right-hand side of the equation, the time derivative, is expressed in terms of the solution function, the forcing term and any other partial derivatives. Because of this, it means we can train the model to learn the relationship between the solution function combined with the forcing term and the time derivative. Putting this together with \lccref{eq:time_step_function}, it is possible to predict the solution function in a future time based on the current solution function and forcing term. The input functions in this case are both the value of the solution function and the forcing term at the current time step. The output function is the time derivative which is computed by rearranging \lccref{eq:time_step_function} resulting in \lccref{eq:time_derivative_time_step_function}.
\begin{equation}
  \dv{f}{t} = \frac{f^{t+1} - f^{t}}{\delta t} \label{eq:time_derivative_time_step_function}
\end{equation}
Perhaps in a more obvious manner, the same is also applied to the WeatherBench 2 collection of ERA5 datasets. The 2-meter temperature is a series of snapshots in time of the distribution of temperature all over the world at 2 meters above the surface. The input function in this case is the temperature distribution at the current time step. The output function is the time derivative of the temperature distribution computed using \lccref{eq:time_derivative_time_step_function}.

@@

The training can proceed to solve \lccref{eq:lssvr_solution}. As the original LSSVR equation was made for regression of a single value of \(\vb{y} \), we constructed an extended the original equation in order to train multiple regression models to predict each coefficient in an output function sample. The assumption of this extension is that the inputs and model hyperparameters such as kernels and the regularization constant \(C \) are the same across individual regressors. Therefore, it follows that the solution vector and target vector can both be transformed into their matrix form with each column representing the individual regressor which results in \lccref{eq:lssvr_solution_vectorized}. This vectorized version is equal to \(m \) separate instances of \lccref{eq:lssvr_solution}. In essence, instead of a vector of scalar sample output values and a vector of model parameters, the extended equation consists of a matrix of rows of vector output values and a matrix of model parameters with each column representing the separate model for each output and row representing the Lagrange multipliers for each sample. The procedure to solve \lccref{eq:lssvr_solution_vectorized} adjusts \lccref{alg:lssvr_training} for the extra dimension spanning the multiple outputs of size \(m\).
\begin{equation} \label{eq:lssvr_solution_vectorized}
  \begin{bmatrix}
    0          & \vb{1}_{n}^{\intercal} \\
    \vb{1}_{n} & \Omega + \frac{I}{C}
  \end{bmatrix}
  \begin{bmatrix}
    b_0           & \cdots & b_m           \\
    \vb{\alpha}_0 & \cdots & \vb{\alpha}_m \\
  \end{bmatrix}
  =
  \begin{bmatrix}
    0        & \cdots & 0        \\
    \vb{y}_0 & \cdots & \vb{y}_m \\
  \end{bmatrix}
\end{equation}

@@

\noindent The utilities module provide common tools that are used when working with the developed software. First, because the basis function used in this study is the Fourier basis, a large proportion of concern is the storage of Fourier basis coefficients. The Fourier series representation that is used in this study is the complex representation as show in \lccref{eq:fourier_series}. However, as previously mentioned, least squares support vector machines is constructed to work with real numbers. In order for LSSVR to learn the relationship between complex numbers, a representation of complex numbers in terms of real numbers is necessary. Since sample features and labels for LSSVR need to be flattened into one dimension which result in a matrix with rows of samples, the representation of complex numbers using real numbers can be done after features and labels have been flattened. An example of such a matrix of two samples with four features each can be seen in \lccref{eq:complex_matrix}. To represent the complex number features as real numbers, once can simply split a single complex number into two real numbers. When this is done to the entire matrix of complex numbers, the result is \lccref{eq:complex_matrix_real_rep}.
\begin{align}
  \begin{bmatrix}
    -1.03+0i & -0.52+0.39i & -0.54+0i & -0.52-0.39i \\
    -0.85+0i & -0.39+0.19i & 0.87+0i  & -0.39-0.19i
  \end{bmatrix}\label{eq:complex_matrix} \\
  \begin{bmatrix}
    -1.03 & 0 & -0.52 & 0.39 & -0.54 & 0 & -0.52 & -0.39 \\
    -0.85 & 0 & -0.39 & 0.19 & 0.87  & 0 & -0.39 & -0.19
  \end{bmatrix}\label{eq:complex_matrix_real_rep}
\end{align}
There are several ways to do this in Pytorch, such as extracting the real and imaginary element components into separate matrices and then interleaving the columns to create the real representation. However, a more performant version can be made by using built-in functions of the library that directly allows views of complex valued tensors as real tensors and vice versa. The \verb|view_as_real| and \verb|view_as_complex| functions allow the conversion between real and complex representations by representing complex valued tensors by adding a dimension of size two for each component of complex numbers. The resulting implementation can be seen in \lccref{fig:complex_matrix_conversion_implementation}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
def to_real_coeff(x: torch.Tensor) -> torch.Tensor:
    return torch.view_as_real(x).flatten(-2)

@@

\begin{table}[H]
  \caption{The configuration of experimental scenarios in this study.} \label{table:experimental_scenarios}
  \centering
  \begin{tabularx}{\textwidth}{ L{0.75} *{3}{L{1.09}} } % compute these using approx barycentric coords (x + 3y = 4, x = 0.8y)
    \toprule
    Configuration                    & Scenario 1                                          & Scenario 2                                                          & Scenario 3                                 \\
    \midrule
    Data Source                      & Synthetic (computed from equations)                 & Synthetic (computed from equations) with three viscosity values     & ERA5 dataset                               \\\addlinespace[0.5em]
    Data Augmentations               & Perturbations of three different noise levels       & Perturbation of 10\% the mean standard deviation of function values &                                            \\\addlinespace[0.5em]
    Data Sample Size                 & 5000 unique functions for each noise level          & 500 unique functions for each viscosity value                       & 7000 time steps                            \\\addlinespace[0.5em]
    Number of Modes                  &                                                     &                                                                     &                                            \\\addlinespace[0.5em]
    Scaling                          & Standard scaling                                    & Standard scaling                                                    & Standard scaling                           \\\addlinespace[0.5em]
    Dataset Split                    & 80\% Training and 20\% Testing for each noise level & 80\% Training and 20\% Testing for each viscosity value             & 80\% Training and 20\% Testing             \\\addlinespace[0.5em]
    Regularization Parameter (\(C\)) & 1.0                                                 & 1.0                                                                 & 1.0                                        \\\addlinespace[0.5em]
    Kernel                           & RBF (\(\sigma = features \times X.var()\))          & RBF (\(\sigma = features \times X.var()\))                          & RBF (\(\sigma = features \times X.var()\)) \\\addlinespace[0.5em]
    Additional Tests                 & Exact solutions                                     & Exact solutions                                                     &                                            \\\addlinespace[0.5em]
    Interpretation                   & p-matrix and correlation image                      & p-matrix and correlation image                                      & p-matrix and correlation image             \\
    \bottomrule
  \end{tabularx}
\end{table}

Each scenario serves a different purpose. The first scenario is constructed to establish whether the model can learn any relationships at all in data that implicitly defines differential equations. This scenario is a proof of concept of whether the model is capable of accurately learning the relationship between derivatives and their anti-derivatives. As the simplest case, it provides a foundation for understanding the model and its behavior. The 5\% noise model is then used to predict the antiderivative of a cosine function with a period of one as shown in \lccref{eq:sine_derivative}. The exact solution is known as \lccref{eq:sine_function} \autocite{abramowitzHandbookMathematicalFunctions1972}. The second goal of this scenario is to determine how noise affects the model's predictions. This is especially important for the inverse problem because of how sensitive the derivative is to changes in the antiderivative. To further solidify how the model fares for inverse problems, this scenario will also assess the performance of a trained model for a very simple inverse problem setup to determine the derivatives of the testing subset from the antiderivatives. This setup will utilize the model as a surrogate forward solver and optimize for a loss function using the inverse function of the SpectralSVR class as described in \lccref{sec:impl_model}.
\begin{align}
  f(x) & = 2\pi cos\left(2\pi x\right)\label{eq:sine_derivative} \\
  f(x) & = sin\left(2\pi x\right)\label{eq:sine_function}
\end{align}

@@

\subsection{Scenario 2}
\noindent The second scenario results are presented in two parts which are the coefficient predictions and function value predictions. The functions are first preprocessed in order to format them into current and next values in time as discussed in \lccref{sec:data_transformation}. The coefficient value predictions are shown in \lccref{table:scenario_2_spectral_metrics}. The RBF kernel parameter computed from the data is once again the same for all versions of the dataset with a value of 10. The testing metrics show that the performance of the model vary with the viscosity value \(\nu \). Starting with the inviscid Burgers' equation, with \(\nu=0\), the model performs relatively well and close to the best result with \(\nu=0.01\). The function value evaluated from the predictions is shown in \lccref{table:scenario_2_function_metrics}. The metrics across the board shows that the function value relative metrics are worse compared to the coefficients. Comparing the absolute metrics for the same table, we see the same general trend that the higher viscosity show the model performing worse.

\begin{table}[H]
  \caption{Performance metrics of coefficient prediction of next time step in scenario 2 by noise level.}\label{table:scenario_2_spectral_metrics}
  \centering
  \begin{tabular}{lrrrrr}
    \toprule
    \(\nu\) & MSE  & RMSE & MAE  & R2   & sMAPE \\
    \midrule
    0.0     & 0.01 & 0.09 & 0.06 & 0.87 & 0.60  \\
    0.01    & 0.01 & 0.08 & 0.06 & 0.90 & 0.56  \\
    0.1     & 0.03 & 0.17 & 0.13 & 0.49 & 1.09  \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[H]
  \caption{Performance metrics of function values evaluated from coefficient prediction of next time step in scenario 2 by noise level.}\label{table:scenario_2_function_metrics}
  \centering
  \begin{tabular}{lrrrrr}
    \toprule
    \(\nu\) & MSE      & RMSE & MAE  & R2   & sMAPE \\
    \midrule
    0.0     & 6.36e-04 & 0.03 & 0.02 & 0.85 & 0.66  \\
    0.01    & 5.19e-04 & 0.02 & 0.02 & 0.88 & 0.62  \\
    0.1     & 2.28e-03 & 0.05 & 0.04 & 0.45 & 1.13  \\
    \bottomrule
  \end{tabular}
\end{table}
